---
title: "Time Series Analysis-GDP"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```


## Basic Time Series Analysis

This part is for basic time series analysis on Quarterly GDP data **(source:IMF)**. It shows quarterly Real Gross Domestic Product seasonally Un-adjusted calculated in domestic currency (value in Million) for the period 2nd Quarter 2004 to 3rd Quarter 2022.

```{r gdp_data1, echo=FALSE, message=FALSE, warning=FALSE, }
library(scales)
library(ggplot2)
library(TSstudio)
library(plotly)
library(forecast)
library(kableExtra)
library(DT)
library(matlib)
library(dplyr)

gdp_data  <- read.csv("C:/Users/91981/Desktop/Uday/Personal_analysis_projects/Git_projects/Showcase_projects/showcase_project_website/data/GDP_IMF_India_quarterly_Q22004_Q32022.csv")

Nominal_gdp_unadjusted_rs <- gdp_data[,c(1,2,3)]

Real_gdp_unadjusted_rs <- gdp_data[,c(1,2,4)]

Real_gdp_unadjusted_rs%>%
  knitr::kable(caption = "Table 1. Quarterly Gross Domestic Product, Real, Undjusted, Domestic Currency 
               (2nd Quarter 2004 to 3rd Quarter 2022)")%>%
  kable_styling(bootstrap_options = "striped",
                latex_options = "striped")

```

There are total 74 rows using `nrow()` formula which means it is equivalent to 74 quarters or  almost 18 years data. Now to apply time series analysis it is important to convert the data in **TS (Time Series)** format. Figure 1 below shows the interactive line chart to see the visual trend in the data, as expected it shows increasing trend with a steep fall due to COVID-19 pandemic.

```{r ts_gdp_data1_plot, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Fig 1. India quarterly Real GDP in Domestic Currency [Rs in Lakh Crores]"}

options(scipen = 100, digits = 4)

ts_gdp_data1 <- with(Real_gdp_unadjusted_rs, ts(NGDP_R_NSA_XDC*(1/1000000), start = c(Year[1], Quarter[1]), freq = 4))

ts_plot(ts_gdp_data1,
        title = "India Quarterly Gross Domestic Product (Real - Undjusted)
        at Domestic Currency [Rs in Lakh Crores]",
        Xtitle = "Period",
        Ytitle = "Rs in Million",
        slider = TRUE,
        color = "blue",
        width = 3,
        line.mode = "lines+markers"
        )


```

**Note:** The variation in data is very high since 2020 (COVID-19 Pandemic starts).

It is best practice to convert it into log form to see the linear trend in the data instead of changing y-axis units.

```{r log_ts_gdp_data1_plot, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Fig 2. India quarterly Real log GDP"}

options(scipen = 100, digits = 4)

ts_gdp_data2 <- with(Real_gdp_unadjusted_rs, ts(log(NGDP_R_NSA_XDC*1000000), start = c(Year[1], Quarter[1]), freq = 4))

ts_plot(ts_gdp_data2,
        title = "India Quarterly log of Real GDP",
        Xtitle = "Period",
        Ytitle = "Log(Real GDP)",
        slider = TRUE,
        color = "green",
        width = 3,
        line.mode = "lines+markers"
        )


```

Now, if we apply linear regression model on the Nominal Un-adjusted quarterly GDP then,

```{r ts_lm_model1}

gdp.linear.model1 = lm(ts_gdp_data2 ~ time(ts_gdp_data2))

summary(gdp.linear.model1)

```
Here, the overall model is statistically significant as the p-value is very small. Both the intercept and slope coefficients are statistically significant but the intercept coefficient does not provide any economic reasoning due to negative value.

```{r basic_lm_graph, fig.cap="Fig 2. Log of India Quarterly Real GDP and linear trend", fig.align='center'}
options(scipen = 100, digits = 4)
plot(ts_gdp_data2,
        main = "Log of India Quarterly Real GDP and 
     Linear Regression Trend",
        xlab = "Period (Quarter)",
        ylab = "Log(Real GDP)",
        col = "blue",
        cex = 2)
abline(gdp.linear.model1, col='red')

```

Figure 2 shows the linear trend and log of quarterly India Real GDP values. One of the important assumption for linear regression is **Residuals Distributed Normally**. It is also to know the residuals are random or not, which is essential for explaining relation in the data.

```{r residual, fig.cap="Fig 3. Histogram of Residuals to Check Normality", fig.align='center'}
options(scipen = 100, digits = 4)
gdp.residuals = resid(gdp.linear.model1)

hist(gdp.residuals, main = "Histogram of Residuals")
lines(density(gdp.residuals))
```

```{r qqplot, fig.cap="Fig 4. Q-Q Plot to check Normality", fig.align='center'}
options(scipen = 100, digits = 4)
qqnorm(gdp.residuals, main = "Normal Q-Q plot", col="darkgrey")
qqline(gdp.residuals,  col="dodgerblue")

```


```{r fitted_vs_residuals, fig.cap="Fig 5. Residuals vs fitted - Homoscedasticity check"}
options(scipen = 100, digits = 4)
plot(fitted(gdp.linear.model1), resid(gdp.linear.model1), col = "grey", xlab = "Fitted", ylab = "Residuals", main = "Fitted vs Residuals")
abline(h=0, col="darkorange")
```

```{r normality_test}

shapiro.test(gdp.residuals)

```

As we can see that the residuals are Non-normally distributed or Skewed distributed both visually and using the test. So, this model can't be used for inferential or predictive purpose.

## Time Series Modeling

As we have seen above that the normal linear regression is not able to predict or make inferences for time series data. The main reason for that is natural correlation from its lag values *(time variables are auto correlated)*.

```{r time_series_modeling_1}
library(astsa)
```

**Stationary time series** is an important concept from time series modelling perspective because time series models build on stationary time series data. Here Weak stationary time series is the one with,

**1. No trend** - we want our time series data to be stationary which means we do not want any systematic change in mean like no trend in the data.  

**2. No Change in Variation** - We do not want a systematic change in variation. No Periodic Fluctuations.

In real world we do have the data as Non-stationary time series data like shown in above **Fig 1**. So we do some transformation if our data is non stationary.


```{r autocorrelation}
options(scipen = 100, digits = 4)
acf(ts_gdp_data2, main = 'Correlogram of Log of Real GDP')

```

To go further for time series modeling, it is important to make it stationary.


```{r detrend_gdp}
options(scipen = 100, digits = 4)

detrend_gdp = diff(ts_gdp_data2)

plot(detrend_gdp, main = 'Detrend log of Real GDP dataset', ylab = 'Log(Real GDP)', xlab = 'Quarter', col = 'blue', lwd = 2)

```

Correlogram after detrend the data,

```{r detrended acf}
options(scipen = 100, digits = 4)
acf(detrend_gdp, main = 'Correlogram of Detrended log of Real GDP')

```



___
**WIP**






---

<div class="footer">Copyright &copy; Uday Khanna Doe 2023</div>